---
title: "Questionnaire"
output: html_document
date: "2024-10-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6)
```

```{r}
install.packages("tidytext")
library(tidytext)
library(tidyverse)
library(readxl)
# install.packages("tidyr")
library(reshape2)
library(pheatmap)
install.packages('RColorBrewer')
library(RColorBrewer)
```

```{r}
theme_set(theme_bw())
```

```{r}
library(readr)
```



# Import the Questionnaire
```{r}
Quest_all <- read_excel("Quest_all.xls")
```


```{r}
Quest_all
```

# order country levels and modify names (medit and alp italy)
```{r}
Quest_all$Country<-factor(Quest_all$Country, levels=c('Sweden', 'Germany', 'Italy (TN)', 'Spain', 'Portugal', 'Italy (RM)'))
levels(Quest_all$Country)<-c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)')
```


## *Professional sector of respondents*

# Split the column with *profess activity*, to include all keywords
```{r}
pro_actvity=
colsplit(Quest_all$`Which of the following categories best suits your professional activity in the river and riparian zone?`, ',', c('a', 'b', 'c', 'd'))


```

# fill the empty cells with NA
```{r}
pro_actvity$b[pro_actvity$b=='']<-NA
pro_actvity$c[pro_actvity$c=='']<-NA
pro_actvity$d[pro_actvity$d=='']<-NA

```


```{r}
pro_actvity=
cbind(Quest_all[,c(1,2,3,16,17,21,22)] , pro_actvity)
pro_actvity$id=c(1:nrow(pro_actvity))

```

# convert to long format to account for multiple profession keywords
```{r}
pro_activity.long=
pro_actvity %>% 
  pivot_longer(c(8,9,10,11), values_to = 'Sector')
```

# trim some leading withe spaces
```{r}
pro_activity.long$Sector=
trimws(pro_activity.long$Sector, 'left')
```


```{r}
names(pro_activity.long)[2]<-'Gender'
names(pro_activity.long)[5]<-'Conservation_Status'

```

```{r}
pro_activity.long$Conservation_Status<-
  factor(pro_activity.long$Conservation_Status, levels = c('High', 'Good', 'Moderate', 'Low'))
```


# check freq of keywords, for contry, sex, etc
```{r}

pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Country)

pro_activity.long %>% 
  #filter(`Gender you identified yourself with` == 'Male') %>% 
  filter(complete.cases(.)) %>% 
  count( Country, Sector)

pro_activity.long %>% 
  #filter(`Gender you identified yourself with` == 'Male') %>% 
  filter(complete.cases(.)) %>% 
  count(Gender,Sector)
```


# Some plots of Professional sector

```{r fig.width=6}
pro_activity.long %>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  filter(Gender!='Prefer not to say') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Gender,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=Gender)+geom_bar(position = 'fill', stat='identity') +ylab(NULL)+
  scale_fill_manual(values=c('orange3', 'steelblue'))+xlab(NULL)+theme(axis.text.x = element_text(angle=45, hjust=1))

pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Sector,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=Sector)+geom_bar(position = 'fill', stat='identity')+ylab(NULL)+
  scale_fill_brewer(palette = 'Paired')+xlab(NULL)+theme(axis.text.x = element_text(angle=45, hjust=1))

pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Conservation_Status,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=Conservation_Status)+geom_bar(position = 'fill', stat='identity')+ylab(NULL)+
  scale_fill_manual(values=c('#2b83ba', '#abdda4', '#fdae61', '#d7191c'))+xlab(NULL)+theme(axis.text.x = element_text(angle=45, hjust=1))


pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(`CAP buffer zones`,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=`CAP buffer zones`)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1))
```


```{r}
pdf('CAP.pdf', w=5, h=4)
pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  #filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(`CAP buffer zones`,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=`CAP buffer zones`)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1))
dev.off()
```




# Check the %responses for CAP, Sex and Country
```{r}
pro_activity.long%>% 
  #filter(complete.cases(.)) %>% 
  #filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(`CAP buffer zones`) %>% 
  mutate(prop=n/sum(n))



pro_activity.long %>% 
  #filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  filter(Gender!='Prefer not to say') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Gender) %>% 
    mutate(prop=n/sum(n))

Quest_all %>% 
  count(Country) %>% 
  mutate(prop=n/sum(n))


```

```{r}
pro_activity.long$Country<-factor(pro_activity.long$Country, levels=c('Sweden', 'Germany', 'Italy (TN)', 'Spain', 'Portugal', 'Italy (RM)'))
```


# export plot conservation status
```{r}

pdf('pie_status.pdf', h=3, w=3)
pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Conservation_Status) %>% 
  mutate(hole=3) %>% 
  mutate(prop=n/sum(n)) %>% 
    ggplot()+aes(x=hole, y=n, fill=Conservation_Status)+geom_col()+coord_polar(theta = 'y')+xlim(c(0.2, 3 + 0.5))+
scale_fill_manual(values=c('#2b83ba', '#abdda4', '#fdae61', '#d7191c'))+theme(legend.position = 'none')+
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())
dev.off()

pdf('plot_status.pdf', w=6, h=4)
pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Conservation_Status,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=Conservation_Status)+geom_bar(position = 'fill', stat='identity')+ylab(NULL)+
  scale_fill_manual(values=c('#2b83ba', '#abdda4', '#fdae61', '#d7191c'))+xlab(NULL)+theme(axis.text.x = element_text(angle=45, hjust=1), axis.text = element_text(size=12))+
  guides(fill=guide_legend(ncol=2))
dev.off()
```

```{r}
%>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Conservation_Status)
```



# Plot of Sectors
```{r}
plot_sector<-
  pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Sector,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=Sector)+geom_bar(position = 'fill', stat='identity')+ylab(NULL)+
  scale_fill_brewer(palette = 'Paired')+xlab(NULL)+theme(axis.text.x = element_text(angle=45, hjust=1))+
  guides(fill=guide_legend(ncol=2))
  

pie_sector<-
pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Sector) %>% 
  mutate(hole=3) %>% 
  mutate(prop=n/sum(n)) %>% 
    ggplot()+aes(x=hole, y=n, fill=Sector)+geom_col()+coord_polar(theta = 'y')+xlim(c(0.2, 3 + 0.5))+
 scale_fill_brewer(palette = 'Paired')+theme(legend.position = 'none')+
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())



```
# Export sector
```{r}
pdf('plot_sector.pdf', w=8, h=5)
plot_sector
dev.off()


pdf('pie_sector.pdf', w=3, h=3)
pie_sector
dev.off()
```

```{r}
Quest_all$Age<-factor(Quest_all$Age, levels=c('18-25', '26-35', '36-45', '46-55', '56-65', '>65', 'NA'))

Quest_all %>% 
    count( Gender) 
```


```{r fig.width=3, fig.height=2.5}

pdf('Age.pdf', w=4, h=3)
Quest_all %>% 
  filter(Age !='NA') %>% 
  filter(Gender != 'Not declared') %>% 
   filter(Gender!='Prefer not to say') %>% 
  count(Age, Gender) %>% 
  ggplot()+aes(x=Age, y=n, fill=Gender)+geom_bar(stat='identity', position='dodge')+theme_classic()+
  scale_fill_manual(values=c('orange2', 'steelblue'))+theme(legend.title=element_blank(), legend.position = c(0.24, 0.8))+ylab(NULL)
dev.off()


```




```{r}
pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender!='Not declared') %>% 
  #filter(`Gender you identified yourself with` == 'Female') %>% 
  count(Conservation_Status,Country) %>% 
  group_by(Country) %>% 
  mutate(freq= n/sum(n)) 
```




# *Non professional activities*

```{r}
Non_pro_actvity=
colsplit(Quest_all$`Non-professional activities`, ',', c('a', 'b', 'c'))

```

#Combine the non-profess activity with additional info
```{r}
Non_pro_actvity=
cbind(Quest_all[,c(1,2,3,16,17,21,22)] , Non_pro_actvity)
Non_pro_actvity$id=c(1:nrow(Non_pro_actvity))

#fill the empty cells with NA
Non_pro_actvity$a[Non_pro_actvity$a=='']<-NA
Non_pro_actvity$b[Non_pro_actvity$b=='']<-NA
Non_pro_actvity$c[Non_pro_actvity$c=='']<-NA

```

# convert to long format to account for multiple non-prof activity
```{r}
Non_pro_activity.long=
Non_pro_actvity %>% 
  pivot_longer(c(8,9,10), values_to = 'Activity')
```

# trim some leading white spaces
```{r}
Non_pro_activity.long$Activity=
trimws(Non_pro_activity.long$Activity, 'left')

Non_pro_activity.long$Activity=
trimws(Non_pro_activity.long$Activity, 'right')

unique(Non_pro_activity.long$Activity)



```

```{r}
Non_pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  count(Activity,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=Activity)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1))

Non_pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Gender=='Female') %>% 
  #filter(Gender=='Male') %>% 
  count(Activity,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=Activity)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1))+ggtitle('Female') +
  scale_fill_manual(values=c('#d7191c', '#fdae61', '#ffffbf', '#a6d96a','#1a9641'))


Non_pro_activity.long%>% 
  filter(complete.cases(.)) %>% 
  #filter(Gender=='Female') %>% 
  filter(Gender=='Male') %>% 
  count(Activity,Country) %>% 
  ggplot()+aes (x=Country,y=n, fill=Activity)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1))+ggtitle('Male')+
  scale_fill_manual(values=c('#d7191c',  '#ffffbf', '#a6d96a','#1a9641'))


```



### Functions of riparian zones

```{r}
Functions<-
  Quest_all[,c(1,2,3,6:10,22)]

Functions$ID=
  seq(1:nrow(Functions))
```

```{r}
Functions.long<-
  Functions %>% 
  pivot_longer(c(4,5,6,7), names_to = 'Perceived_functions')

Functions.long$value=
  factor(Functions.long$value, levels = c('very important', 'important', 'slightly important'))

Functions.long$value

Functions.long$Country<-factor(Functions.long$Country, levels=c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)'))

```

# Functions
```{r fig.height=5,fig.width=6}


Functions.long%>% 
  filter(complete.cases(.)) %>% 
  count(Perceived_functions, value, Country) %>% 
  ggplot()+aes (x=Perceived_functions, y=n, fill=value)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+ylab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1))+facet_wrap(~Country)+scale_fill_manual(values=c('darkred', 'orange', 'gold'))+
   coord_flip()+theme(legend.position = 'top')


pdf('Functions.pdf', h=6, w=5)
Functions.long%>% 
  filter(complete.cases(.)) %>% 
  count(Perceived_functions, value, Country) %>% 
  ggplot()+aes (x=Perceived_functions, y=n, fill=value)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+ylab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1))+facet_wrap(~Country)+scale_fill_manual(values=c('darkred', 'orange', 'gold'))+
   coord_flip()+theme(legend.position = 'top')
dev.off()

```
# Chi square test on functions
#*temp regulation not sig differences*


```{r}
temp_regul<-
Functions.long %>% 
  filter(Perceived_functions=='Temperature regulation')

csqtest_temp_regul<-
chisq.test(temp_regul$value, temp_regul$Country)


pheatmap(csqtest_temp_regul$residuals,
         cluster_rows = F, cluster_cols = F,display_numbers = T, fontsize_number = 10)

```

#*flood protection is sig different*
```{r}
flood_prot<-
  Functions.long %>% 
  filter(Perceived_functions=='Flood protection')

csqtest_flood_prot<-
  chisq.test(flood_prot$value, flood_prot$Country)

pheatmap(csqtest_flood_prot$residuals,
         cluster_rows = F, cluster_cols = F,display_numbers = T, fontsize_number = 10, main="Flood protection")



```




# Stressors

Shorten column name
```{r}
names(Quest_all)[14]<-'Land use alteration (agriculture)'
names(Quest_all)[13]<-'Invasive species & pests'

```


# Create df on stressors
```{r}
Stressors<-
  Quest_all[,c(1,2,3,11:15,22)]

Stressors$ID=
  seq(1:nrow(Stressors))
```
# long format of stressors
```{r}
Stressors.long<-
 Stressors %>% 
  pivot_longer(c(4,5,6,7,8), names_to = 'Perceived_stressor')

Stressors.long$value=
  factor(Stressors.long$value, levels = c('very important', 'important', 'slightly important'))

Stressors.long$value
```

```{r}
Stressors.long$Country<-factor(Stressors.long$Country, levels=c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)'))
```


# plots of stressors
```{r fig.height=5, fig.width=7}
Stressors.long%>% 
  filter(complete.cases(.)) %>% 
  count(Perceived_stressor, value, Country) %>% 
  ggplot()+aes (x=Perceived_stressor, y=n, fill=value)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+ylab(NULL)+
 theme(axis.text.x = element_text(angle=50, hjust=1))+facet_wrap(~Country)+scale_fill_manual(values=c('darkred', 'orange', 'gold'))+
  coord_flip()+theme(legend.position = 'top')

pdf('Stressors.pdf', h=6, w=5)
Stressors.long%>% 
  filter(complete.cases(.)) %>% 
  count(Perceived_stressor, value, Country) %>% 
  ggplot()+aes (x=Perceived_stressor, y=n, fill=value)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+ylab(NULL)+
 theme(axis.text.x = element_text(angle=50, hjust=1))+facet_wrap(~Country)+scale_fill_manual(values=c('darkred', 'orange', 'gold'))+
  coord_flip()+theme(legend.position = 'top')
dev.off()


```


```{r}
Stressors.long%>% 
  filter(complete.cases(.)) %>% 
  count(Perceived_stressor, value) %>% 
  ggplot()+aes (x=Perceived_stressor, y=n, fill=value)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+ylab(NULL)+
 theme(axis.text.x = element_text(angle=50, hjust=1))+scale_fill_manual(values=c('darkred', 'orange', 'gold'))+
  coord_flip()+theme(legend.position = 'top')
```




# plot of stressors for younger stakeholders
```{r}
Stressors.long%>% 
  filter(complete.cases(.)) %>% 
  filter(Age %in% c('18-25', '26-35', '36-45'))%>%
  count(Perceived_stressor, value, Country) %>% 
  ggplot()+aes (x=Perceived_stressor, y=n, fill=value)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+ylab(NULL)+
 theme(axis.text.x = element_text(angle=50, hjust=1))+facet_wrap(~Country)+scale_fill_manual(values=c('darkred', 'orange', 'gold'))+
  coord_flip()+theme(legend.position = 'top')
```




# Pollution test
```{r}
tmp2=
Stressors.long %>% filter(Perceived_stressor=='Pollution') 
  

# these two commands are the same as 
csqtest_pollution=
chisq.test(table(tmp2$value, tmp2$Country))
# or
chisq.test(tmp2$value, tmp2$Country)


csqtest_pollution$observed
csqtest_pollution$expected
csqtest_pollution$residuals

# display the residuals
pheatmap(csqtest_pollution$residuals,
         cluster_rows = F, cluster_cols = F,display_numbers = T, fontsize_number = 10, main='Pollution')


# contribution to overall chi stats
pollut.contrib=
  100*
  ((csqtest_pollution$observed- csqtest_pollution$expected)^2 /csqtest_pollution$expected)/
  csqtest_pollution$statistic



# display the contribtionn
pheatmap(pollut.contrib,
         display_numbers = T,
         cluster_rows = F,
         cluster_cols = F)  

```
# invasive spp threat test
```{r}
tmp2=
Stressors.long %>% filter(Perceived_stressor=='Invasive species & pests') 
  

# these two commands are the same as 
csqtest_invasive=
chisq.test(table(tmp2$value, tmp2$Country))
# or
chisq.test(tmp2$value, tmp2$Country)


csqtest_invasive$observed
csqtest_invasive$expected
csqtest_invasive$residuals

# display the residuals
pheatmap(csqtest_invasive$residuals,
         cluster_rows = F, cluster_cols = F, display_numbers = T, fontsize_number = 10, main='Invasive spp')

# contribution to overall chi stats
pollut.contrib=
  100*
  ((csqtest_pollution$observed- csqtest_pollution$expected)^2 /csqtest_pollution$expected)/
  csqtest_pollution$statistic



# display the contribtionn
pheatmap(pollut.contrib,
         display_numbers = T,
         cluster_rows = F,
         cluster_cols = F)  
```



## Hot to extract proportion of each answer from each country.
```{r}
# here you get the relative freq of each stressor
Stressors.long%>% 
  filter(complete.cases(.)) %>% 
  #filter(Country =='Sweden') %>% 
  count(Perceived_stressor, value, Country) %>% 
  group_by( Country, Perceived_stressor) %>%
  mutate(freq= n/sum(n)) 

# here you get the relative freq of each individual response
Stressors.long%>% 
  filter(complete.cases(.)) %>% 
  #filter(Country =='Sweden') %>% 
  count(Perceived_stressor, value, Country) %>% 
  group_by(Country) %>% 
  mutate(tot_n=sum(n)) %>% 
  mutate(rel_n=n/tot_n)

```
 


###############
# emblematic species    
```{r}
Species<-
colsplit(Quest_all$`Emblematic species`, '-', c('a', 'b', 'c','d'))

Species$ID=
  seq(1:nrow(Species))

Species<-
  cbind(Quest_all[,c(1,2,3,22)], Species)

Species$b[Species$b=='']<-NA
Species$c[Species$c=='']<-NA
Species$d[Species$d=='']<-NA


```

# Long species df
```{r}
Species.long=
Species %>% 
  pivot_longer(c(5:8), values_to = 'Emblematic species')
```

# time species names
```{r}
Species.long$`Emblematic species`=
trimws(Species.long$`Emblematic species`, 'left')

Species.long$`Emblematic species`=
trimws(Species.long$`Emblematic species`, 'right')
```

# remove NA and make species as factor to sort them
```{r}
Species.long <-
  Species.long %>% 
  filter(complete.cases(.))


Species.long$`Emblematic species`<-
  factor(Species.long$`Emblematic species`)


```




# Sorting species by freq
# create species vector for Worlde
```{r}
species.sorted=
sort(table(Species.long$`Emblematic species`), decreasing = T)

species.sorted<-
  as.data.frame(species.sorted)


write.csv(species.sorted, 'species.sorted.csv')


```

#WORDLE
```{r}
install.packages("wordcloud")
library(wordcloud)

species.sorted_noukn=species.sorted[-3,]

wordcloud(words = species.sorted_noukn$Var1, freq = species.sorted_noukn$Freq, min.freq = 2, max.words=55, random.order=FALSE, 
          rot.per = 0, fixed.asp = F)

```


# rank species according to overall freq
```{r}
Species.long$spp_rank=
  species.sorted$Freq[match(Species.long$`Emblematic species`, species.sorted$Var1)]


species.sorted$Var1<-factor(species.sorted$Var1)
species.sorted$Var1


Species.long$`Emblematic species`<-
  factor(Species.long$`Emblematic species`)

levels(species.sorted$Var1)


Species.long$`Emblematic species`<-
  factor(Species.long$`Emblematic species`, levels=levels(species.sorted$Var1))

Species.long$`Emblematic species`

```

```{r}
Species.long$Country<-factor(Species.long$Country, levels=c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)'))
```

```{r}
install.packages('stringr')
library(stringr)
```

# Rename and homogenize Bat and Bats (all BATS now)
```{r}
tmp=
Species.long %>% 
  filter(`Emblematic species` =='Bat')


which (Species.long$`Emblematic species` == "Bat")
# 568 579 586

Species.long$`Emblematic species`[568]<-'Bats'
Species.long$`Emblematic species`[579]<-'Bats'
Species.long$`Emblematic species`[586]<-'Bats'


```

#Sort the species again, after the name change in BATS
```{r}
species.sorted=
sort(table(Species.long$`Emblematic species`), decreasing = T)

species.sorted<-
  as.data.frame(species.sorted)

# reworder factor in data
Species.long$`Emblematic species`<-
   factor(Species.long$`Emblematic species`, levels=levels(species.sorted$Var1))

#write.csv(species.sorted, 'species.sorted.csv')


```


```{r fig.width=7, fig.height=6}
Species.long %>% 
  filter(complete.cases(.)) %>% 
  filter(`Emblematic species` %in% species.sorted$Var1[1:50]) %>% 
 # arrange(desc(spp_rank)) %>%
  count(`Emblematic species`,Country) %>% 
  ggplot()+aes(x=`Emblematic species`, y=n, fill=Country)+geom_col()+xlab(NULL)+ylab(NULL)+
  theme(axis.text.x = element_text(angle=90, size=7))+scale_fill_manual(values=c('#1b5566','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02'))+
  coord_flip()+theme_classic()+guides(fill=guide_legend(ncol=3))+theme(legend.position = c(0.5,0.92))


Species.long %>% 
  filter(complete.cases(.)) %>% 
  filter(`Emblematic species` %in% species.sorted$Var1[1:50]) %>% 
  count(Country, `Emblematic species`) %>% 
  ggplot()+aes(x=`Emblematic species`, y=n)+geom_col()+theme(axis.text.x = element_text(angle=90, size=7))+xlab(NULL)+ylab(NULL)+
  facet_wrap(~Country, scale='free')

## Excluding the I dont know response
Species.long %>% 
  filter(complete.cases(.)) %>% 
  filter(`Emblematic species` %in% species.sorted_noukn$Var1[1:50]) %>% 
  #arrange(desc(spp_rank)) %>%
  count(`Emblematic species`,Country) %>% 
  ggplot()+aes(x=`Emblematic species`, y=n, fill=Country)+geom_col()+xlab(NULL)+ylab(NULL)+
  theme(axis.text.x = element_text(angle=90, size=7))+scale_fill_manual(values=c('#1b5566','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02'))+
  coord_flip()+theme_classic()+guides(fill=guide_legend(ncol=3))+theme(legend.position = c(0.5,0.92))

Species.long %>% 
  filter(complete.cases(.)) %>% 
  filter(`Emblematic species` %in% species.sorted_noukn$Var1[1:50]) %>% 
  count(Country, `Emblematic species`) %>% 
  ggplot()+aes(x=`Emblematic species`, y=n)+geom_col()+theme(axis.text.x = element_text(angle=90, size=7))+xlab(NULL)+ylab(NULL)+
  facet_wrap(~Country, scale='free')



```
# Export the freq of emblematic spp
```{r}
pdf('Spp_freq.pdf',h=7, w=7)
Species.long %>% 
  filter(complete.cases(.)) %>% 
  filter(`Emblematic species` %in% species.sorted$Var1[1:50]) %>% 
  #arrange(desc(spp_rank)) %>%
  count(`Emblematic species`,Country) %>% 
  ggplot()+aes(x=`Emblematic species`, y=n, fill=Country)+geom_col()+xlab(NULL)+ylab('Number of responses')+
  theme(axis.text.x = element_text(angle=90, size=7))+scale_fill_manual(values=c('#1b5566','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02'))+
  coord_flip()+theme_classic()+guides(fill=guide_legend(ncol=3))+theme(legend.position = c(0.5,0.92))

dev.off()


```


```{r}
Species.long %>% 
  filter(complete.cases(.)) %>% 
  #filter(`Emblematic species` %in% species.sorted$Var1[1:50]) %>% 
  #arrange(desc(spp_rank)) %>%
  group_by(Country) %>% 
  count(`Emblematic species`) %>% 
  mutate(tot=n()) %>% 
  mutate(percent=n/tot) %>% 
  filter(`Emblematic species` == "I don't know" ) %>% 
  select(percent) %>% summary()

```






# Export the freq of emblematic spp (w/o I dont know answers)
```{r}
pdf('Spp_freq_nounkn.pdf',h=7, w=7)
Species.long %>% 
  filter(complete.cases(.)) %>% 
  filter(`Emblematic species` %in% species.sorted_noukn$Var1[1:50]) %>% 
  #arrange(desc(spp_rank)) %>%
  count(`Emblematic species`,Country) %>% 
  ggplot()+aes(x=`Emblematic species`, y=n, fill=Country)+geom_col()+xlab(NULL)+ylab(NULL)+
  theme(axis.text.x = element_text(angle=90, size=7))+scale_fill_manual(values=c('#1b5566','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02'))+
  coord_flip()+theme_classic()+guides(fill=guide_legend(ncol=3))+theme(legend.position = c(0.5,0.92))

dev.off()
```




## emblematic species composition, as PCoA
```{r}
Species.composition=
Species.long %>% 
  count(Country, `Emblematic species`) %>% 
  pivot_wider(names_from = `Emblematic species`, values_from = n)
```

```{r}
library(vegan)
```


```{r}
Species.composition[is.na(Species.composition)]<-0


dist.bray=vegdist(Species.composition[,-1], 'bray')

dist.bray=vegdist(Species.composition[,c(2:50)], 'bray')


dist.bray_nounkn=vegdist(Species.composition[,-c(1,4)], 'bray')


```

## PCoA with APE 
```{r}
library(ape)

pcoa2<-
pcoa(dist.bray)
biplot.pcoa(pcoa2, Species.composition[,c(2:5)], cex=0.1)


pcoa2_nouknw<-
pcoa(dist.bray_nounkn)
biplot.pcoa(pcoa2_nouknw, Species.composition[,c(4:12)], cex=0.1)


```



# Plot of the PCoA (species composition)

```{r}
library(ggrepel)
```


# PCOA cmdscale
*removing idontknow answer give same results*
```{r}


pcoa1<-cmdscale(dist.bray, eig=T)
pcoa1$eig

pcoa1=as.data.frame(
cmdscale(dist.bray))

pcoa1$Country=Species.composition$Country
pcoa1$Country<-factor(pcoa1$Country, levels=c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)'))

pcoa1 %>% 
  ggplot()+aes(V1, V2, col=Country, label=Country)+geom_point()+scale_color_manual(values=c('#1b5566','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02'))+
  xlim(c(-0.3,0.6))+geom_text_repel()



pcoa1_nounkn<-cmdscale(dist.bray_nounkn, eig=T)
pcoa1_nounkn<-as.data.frame(cmdscale(dist.bray_nounkn))
pcoa1_nounkn$Country=Species.composition$Country
pcoa1_nounkn$Country<-factor(pcoa1$Country, levels=c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)'))

pcoa1_nounkn %>% 
  ggplot()+aes(V1, V2, col=Country, label=Country)+geom_point()+scale_color_manual(values=c('#1b5566','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02'))+
  geom_text_repel()

```

# extract spp scores    *notworking*
```{r}
install.packages('BiodiversityR')
library(BiodiversityR)
install.packages('tcltk')
library(tcltk)

pcoa1_sppscore<-
add.spec.scores(pcoa1, Species.composition[,-1], method="cor.scores", multi=1, Rscale=F, scaling="1")
```


# export PCoA plot
```{r fig.width=5}
pcoa1 %>% 
  ggplot()+aes(V1, V2, col=Country, label=Country)+geom_point()+scale_color_manual(values=c('#1b5566','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02'))+
  xlim(c(-0.3,0.6))+geom_text_repel()+xlab('PCoA1 40%')+ylab('PCoA2 21%')+theme(legend.position = 'none')+theme_bw()

pdf('PCoA1.pdf', w=3, h=3)
pcoa1 %>% 
  ggplot()+aes(V1, V2, col=Country, label=Country)+geom_point()+scale_color_manual(values=c('#1b5566','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02'))+
  xlim(c(-0.3,0.6))+geom_text_repel()+xlab('PCoA1 40%')+ylab('PCoA2 21%')+theme(legend.position = 'none')
dev.off()

```




########
## management actions


```{r}

Quest_all <- read_excel("Quest_all.xls")
Quest_all$Country<-factor(Quest_all$Country, levels=c('Sweden', 'Germany', 'Italy (TN)', 'Spain', 'Portugal', 'Italy (RM)'))
levels(Quest_all$Country)<-c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)')
```

 
```{r}
Management<-
colsplit(Quest_all$`Management Action - Category`, '-', c('a', 'b', 'c','d'))

Management$ID=
  seq(1:nrow(Management))

Management<-
  cbind(Quest_all[,c(1,2,3,22)], Management)

Management$b[Management$b=='']<-NA
Management$c[Management$c=='']<-NA
Management$d[Management$d=='']<-NA


```



# Long management df
```{r}
Management.long=
Management %>% 
  pivot_longer(c(5:8), values_to = 'Management action')

Management.long$Country<-
factor(Management.long$Country, levels=c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)'))


```

# Trim names
```{r}
Management.long$`Management action`=
trimws(Management.long$`Management action`, 'left')

Management.long$`Management action`=
trimws(Management.long$`Management action`, 'right')
```

```{r}
Management.long$`Management action`[1905]<-'Awareness & Education'
which(Management.long$`Management action`=='Barriers')
Management.long$`Management action`[990]<-'Flow & Barriers'
Management.long$`Management action`[c(1585, 2010)]<-'Flow & Barriers'

which(Management.long$`Management action`=="I don't know")

```


```{r}
table(Management.long$`Management action`, Management.long$Country)


Management.long$`Management action`<-factor(Management.long$`Management action`,levels=c(
'Buffer', 'Vegetation', 'Banks', 'Flow & Barriers',  'Floodplains & Meandering','Let it run wild',
'Protection', 'Restoration', 'Invasives',  'Pollution', 'Sustainable use & Hydropower',
'Monitoring & surveillance', 'Awareness & Education', "I don't know"))

Management.long$`Management action`

```

# Plot management needs + piechart tot
```{r fig.height=5}
plot_managment<-
Management.long%>% 
  filter(complete.cases(.)) %>% 
  count(`Management action`, Country) %>% 
  ggplot()+aes (x=Country, y=n, fill=`Management action`)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1)) +coord_flip()+ylab(NULL)+
  
  scale_fill_manual(values = c(brewer.pal(name="Spectral", n =5), brewer.pal(name="BrBG", n = 8), 'grey80'))+guides(fill=guide_legend(ncol=2))


pie_management<-
Management.long%>% 
  filter(complete.cases(.)) %>% 
    count(`Management action`) %>% 
  mutate(hole=3) %>% 
  mutate(prop=n/sum(n)) %>% 
    ggplot()+aes(x=hole, y=n, fill=`Management action`)+geom_col()+coord_polar(theta = 'y')+xlim(c(0.2, 3 + 0.5))+
  scale_fill_manual(values = c(brewer.pal(name="Spectral", n =5), brewer.pal(name="BrBG", n = 8), 'grey80'))+theme(legend.position = 'none')+
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

```


```{r}
pdf('plot_management.pdf', h=5, w=8)
plot_managment
dev.off()

pdf('pie_management.pdf', h=3, w=3)
pie_management
dev.off()



```
# Plot management with legend in 2 cols to add the CAP plot
```{r}
pdf('plot_management_narrow.pdf', h=5, w=6)
Management.long%>% 
  filter(complete.cases(.)) %>% 
  count(`Management action`, Country) %>% 
  ggplot()+aes (x=Country, y=n, fill=`Management action`)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1), axis.text.y = element_text(size=9)) +coord_flip()+ylab(NULL)+
  
  scale_fill_manual(values = c(brewer.pal(name="Spectral", n =5), brewer.pal(name="BrBG", n = 8), 'grey80'))+guides(fill=guide_legend(ncol=1))+theme(legend.text = element_text(size=8.5))
dev.off()
```
 
 # plot of management needs for younger folks
```{r}
Management.long%>% 
  filter(complete.cases(.)) %>% 
   filter(Age %in% c('18-25', '26-35', '36-45'))%>%
  count(`Management action`, Country) %>% 
  ggplot()+aes (x=Country, y=n, fill=`Management action`)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1), axis.text.y = element_text(size=9)) +coord_flip()+ylab(NULL)+
  
  scale_fill_manual(values = c(brewer.pal(name="Spectral", n =5), brewer.pal(name="BrBG", n = 8), 'grey80'))+guides(fill=guide_legend(ncol=1))+theme(legend.text = element_text(size=8.5))
```





# Chi square test on management action
```{r}
Management.long_noNA<-
  Management.long %>% filter(complete.cases(.))


csqtest_management<-
chisq.test(Management.long_noNA$`Management action`, Management.long_noNA$Country)


pheatmap(csqtest_management$residuals,
         cluster_rows = F, cluster_cols = F, display_numbers = T, fontsize_number = 10, main='Management actions')+theme(legend.position = 'none')
```


# Test plotting with ggalluvium
```{r}
install.packages('ggalluvial')
library(ggalluvial)
```

```{r}
Management.long%>% 
  filter(complete.cases(.)) %>% 
  count(`Management action`, Country) %>% 
  ggplot()+aes(axis1=Country, axis2=`Management action`, y=n)+geom_alluvium(aes(fill=Country))+
  geom_stratum()+geom_text(stat = "stratum",aes(label = after_stat(stratum)))
```




## How should do more?

```{r}

Quest_all <- read_excel("Quest_all.xls")


Quest_all$Country<-factor(Quest_all$Country, levels=c('Sweden', 'Germany', 'Italy (TN)', 'Spain', 'Portugal', 'Italy (RM)'))
levels(Quest_all$Country)<-c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)')
```

 
```{r}
Do_more<-
colsplit(Quest_all$`Who should do more?`, ',', c('a', 'b', 'c','d','e','f','g','h'))

Do_more$ID=
  seq(1:nrow(Do_more))

Do_more<-
  cbind(Quest_all[,c(1,2,3,22)], Do_more)

Do_more$a[Do_more$a=='']<-NA
Do_more$b[Do_more$b=='']<-NA
Do_more$c[Do_more$c=='']<-NA
Do_more$d[Do_more$d=='']<-NA
Do_more$e[Do_more$e=='']<-NA
Do_more$f[Do_more$f=='']<-NA
Do_more$g[Do_more$g=='']<-NA
Do_more$h[Do_more$h=='']<-NA

```


# Long management df
```{r}
Do_more.long=
Do_more %>% 
  pivot_longer(c(5:12), values_to = 'Should_do_more')

Do_more.long$Country<-
factor(Do_more.long$Country, levels=c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)'))
levels(Do_more.long$Country)<-c('Sweden', 'Germany', 'Italy (Alp)', 'Spain', 'Portugal', 'Italy (Med)')
```

# Trim names
```{r}
Do_more.long$Should_do_more=
trimws(Do_more.long$Should_do_more, 'left')

Do_more.long$Should_do_more=
trimws(Do_more.long$Should_do_more, 'right')
```

```{r}
table(Do_more.long$Should_do_more)
table(Do_more.long$Should_do_more, Do_more.long$Country)

which(Do_more.long$Should_do_more=='Water managers')

```

```{r}
plot_domore<-
Do_more.long%>% 
  filter(complete.cases(.)) %>% 
  count(Should_do_more, Country) %>% 
  ggplot()+aes (x=Country, y=n, fill=Should_do_more)+geom_bar(position = 'fill', stat='identity')+xlab(NULL)+
 theme(axis.text.x = element_text(angle=45, hjust=1)) +coord_flip()+ylab(NULL)+
  
  scale_fill_manual(values = c(brewer.pal(name='Dark2', n =7), 'darkblue'))+guides(fill=guide_legend(ncol=2))


pie_domore<-
Do_more.long %>% 
  filter(complete.cases(.)) %>% 
    count(Should_do_more) %>% 
  mutate(hole=3) %>% 
  mutate(prop=n/sum(n)) %>% 
    ggplot()+aes(x=hole, y=n, fill=Should_do_more)+geom_col()+coord_polar(theta = 'y')+xlim(c(0.2, 3 + 0.5))+
  scale_fill_manual(values = c(brewer.pal(name='Dark2', n =7), 'darkblue')) +theme(legend.position = 'none')+
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())



```

```{r}
pdf('plot_domore.pdf', w=8, h=5)
plot_domore
dev.off()

pdf('pie_domore.pdf', w=3,h=3)
pie_domore
dev.off()
```



# Chi sqre test on 'Responsibilities'


```{r fig.width=6}
Do_more.long_noNA<-
  Do_more.long %>% filter(complete.cases(.))

csqtest_do_more<-
chisq.test(Do_more.long_noNA$Should_do_more, Do_more.long_noNA$Country)

pheatmap(csqtest_do_more$residuals,
         cluster_rows = F, cluster_cols = F, display_numbers = T, fontsize_number = 8, main='Responsibilities')
```

```{r}
Do_more_nonac.long_noNA<-
  Do_more_nonac.long %>% filter(complete.cases(.))

csqtest_do_more<-
chisq.test(Do_more_nonac.long_noNA$Should_do_more, Do_more_nonac.long_noNA$Country)


pheatmap(csqtest_do_more$residuals,
         cluster_rows = F, cluster_cols = F, display_numbers = T, fontsize_number = 10, main='Responsibilities')
```
# Prova ggluvium
```{r}
Do_more.long %>% 
  filter(complete.cases(.)) %>% 
  count(Should_do_more, Country) %>% 
  ggplot()+aes(axis1=Country, axis2=Should_do_more, y=n)+geom_alluvium(aes(fill=Country))+
  geom_stratum()+geom_text(stat = "stratum",aes(label=after_stat(stratum)))+
  scale_fill_manual(values=c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e','#e6ab02')) 

```



